---
title: "Excersice Sheet 7"
fontsize: 11pt
header-includes:
- \usepackage[german]{babel}
- \usepackage{caption}
output:
  html_document: default
  pdf_document: default
highlight: tango
fig_caption: yes
---

\captionsetup[table]{labelformat=empty}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

A broker wants to use linear regression to find out which factors have a large influence on the price of a property. For this purpose, the variables described in Table 1 are given for the last 88 sales in the broker's region.

Table: Table 1 House price record

Variabel   | Description
---------  | -------------
`price`    | house price ($\times$ 1,000 EUR)
`bdrms`    | number bedrooms
`lotsize`  | parking area (m$^2$)
`sqrm`     | house area (m$^2$)
`country`  | $==1$ when in country house style
`lprice`   | `log(price)`
`llotsize` | `log(lotsize)`
`lsqrm`    | `log(sqrm)`

1.  Create a linear regression model with `price` as dependent variable and `bdrms`, `lotsize`, `sqrm` und `country`  as independent variables. 
    a) Determine the regression coefficients and  $p$-values of the dependent variable and compare their influence within the model on the predicted value for  `price`. 
    b) Determine how much variance of the dependent variable is explained. 
    c) Check the residuals (graphically) for normal distribution and homoskedasticity.
    
```{r}
# Solution for Task 1...
height<-read.csv("C:/Users/atray/OneDrive/Documents/R/hprice.csv")
lmprice<- lm(price~bdrms+lotsize+sqrm+country, data = height )
summary(lmprice)
anova(lmprice)
plot(height, pch=16, col="blue")
abline(lmprice)


hist(residuals(lmprice))
boxplot(residuals(lmprice))

#par(mfrow=c(2,2))
plot(lmprice)

```

2. Given be the linear regression model from task 1.
    a) Create a scatterplot to display the relationship between the predicted value for `price` and the residual size.
    b) For some houses, the price forecast of the broker model is more than EUR 100,000 off. Highlight houses with a residual size of more than 100 or less than 100. What could be the reasons for high model inaccuracies?
    c) Can the $R^2$-value be increased by using a linear transformation of one of the independent variables?

```{r}
# Solution for Task 2...
library(dplyr)
library(stringr)
plot(height$price,lmprice$residuals,pch=16, col="red")
residuals<-data.frame(residuals(lmprice))
fitted <- data.frame(fitted(lmprice))
data <- data.frame(fitted, residuals)


outliers <- rbind(filter(data, data$residuals.lmprice. < -100), filter(data, data$residuals.lmprice. >100))
plot(x=data$fitted.lmprice., y=data$residuals.lmprice.) +
points(x=outliers$fitted.lmprice, y=outliers$residuals.lmprice, col = "red", pch=10)

height$sqrtbed <- log(height$lotsize)

pricebed <- lm(price ~ bdrms, height)
plot(x=height$bdrms, y=height$price)+
abline(pricebed, col='red')

pricesqrtbed <- lm(price ~ sqrtbed, height)
plot(x=height$sqrtbed, y=height$price)+
abline(pricesqrtbed, col='red')

coefficients(pricebed) 
coefficients(lmprice) 
coefficients(pricesqrtbed)
summary(pricebed) 
summary(lmprice)
summary(pricesqrtbed)
```

3. Graphically display the relationship between `bdrms` and `price`. Check whether this relationship is also reflected in the regression model from Task 1. Create a regression model with `bdrms` as the only independent variable. Compare the regression coefficients with those of the model from Task 1 and interpret the differences. 

```{r}
# Solution for Task 3...
x1<- height$bdrms
y1<- height$price
plot(x1, y1)
lmbdrms<-lm(price~bdrms, data = height)
summary(lmbdrms)
```
